Reaction-Advection-Diffusion Models of Air Pollution

### 2024-04-28 paper.tex
6.5

\documentclass{article}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{physics}
\usepackage{tikz}
\title{Reaction-Advection-Diffusion Models of Air Pollution}
\author{Daniel Henricks}

@@

We will now investigate a puff model; such a model describes fluid elements moving in space. We define a fluid element
to be a volume of air in which all particles move with the same velocity. The most common application of a puff model is diffusion from a 
smokestack, something that is essential for understanding how air pollution diffusion. We define the mass balance equation for this model
to be

@@

\begin{itemize}
    \item Mean velocity ($u$)
    \item Concentration ($C$)
    \item Source term ($S$)
    \item Lateral eddy exchange constant ($K_y$)
    \item Vertical eddy exchange constant ($K_z$)
 \end{itemize}

>>>>

\documentclass{article}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{physics}
\usepackage{tikz}
\title{Reaction-Advection-Diffusion Models of Air Pollution}
\author{Daniel Henricks}

@@

We will now investigate a puff model; such a model describes fluid elements moving in space. Note that the puff model
is an example of a Lagrangian model (as mentioned above). We define a fluid element
to be a volume of air in which all particles move with the same velocity. The most common application of a puff model is diffusion from a 
smokestack, something that is essential for understanding how air pollution diffusion. We define the mass balance equation for this model
to be

@@

\begin{itemize}
    \item Mean wind velocity ($u$)
    \item Concentration ($C$)
    \item Source term ($S$)
    \item Lateral eddy exchange constant ($K_y$)
    \item Vertical eddy exchange constant ($K_z$)
 \end{itemize}

This model was further improved by Buske et. al (2012). In this paper, the initial equation was revised to break the concentration
into a three-dimensional problem as seen below:

\begin{equation}
    u\frac{\partial c}{\partial x} + v\frac{\partial c}{\partial y} + w \frac{\partial c}{\partial z} 
    = \frac{\partial}{\partial x} (K_x \frac{\partial c}{\partial x}) + \frac{\partial}{\partial y} (K_y \frac{\partial c}{\partial y}) + \frac{\partial}{\partial z} (K_z \frac{\partial c}{\partial z})
\end{equation}

This equation essentially states that the mean wind vector can be broken up into its components and that the suspended material must be conserved.
We limit the domains of $x$, $y$, and $z$ to $0 < x < L_x$, $0 < y < L_y$, and $0 < z < h$, where $h$ is the height of the atmospheric boundary layer. 
We also impose a value of zero for the eddy diffusivities at the edges of the domain. Since the solution to the three-dimensional problem can be hard to find,
we want to find a way to transform this equation to a two-dimensional problem. To do this, we need to transform the problem to one with variables $x$ and $z$:

First, we create a complete set of orthogonal basis functions in the y-direction over the y-domain. Define

$$
    Y_m(y) = cos(\lambda_m y)
$$
and 
$$
    \lambda_m = \frac{m\pi}{L_y}, m \in \mathbb{N}^{+}
$$

From doing this, we have created a way for any function $f(y)$ to be represented as some linear combination of these eigenfunctions:

$$
    f(y) = \sum_{m=0}^{\infty} c_m Y_m(y)
$$

We then plug in our unknown coefficient of two variables to get equation (4) in Buske (2012):

\begin{equation}
    f(y) = \sum_{m=0}^{\infty} c_m(x, z) Y_m(y)
\end{equation}

To find the value of this coefficient, we will first define some integrals. Let $Y_n(y)$ also be a orthogonal eigenfunction. Then we define

$$
    \int_0^{L_y} Y_m(y)Y_n(y) dy = \alpha_{mn}
$$

$$
    \int_0^{L_y} Y_m'(y)Y_n(y) dy = \beta_{mn}
$$

$$
    \int_0^{L_y} K_y Y_m(y)Y_n(y) dy = \gamma_{mn}
$$

$$
    \int_0^{L_y} K'_y Y'_m(y)Y_n(y) dy = \eta{mn}
$$

This will help simplify our derivation for finding the unknown constant $c_m(x, z)$.

We first take the chain rule of each of the terms in the diffusion equation and plug in our new equation for the concentration:
$$
    \frac{\partial}{\partial x} [K_x \frac{\partial c(x, y, z)}{\partial x}] = [\frac{\partial}{\partial x} (K_x \frac{\partial c(x, z)}{\partial x}) + K_x \frac{\partial^2 c(x, z)}{\partial x^2}]Y_m(y)
$$

The same process is done for the y-derivative, yielding

$$
    K_y' c_m(x, z) Y_m'(y) + K_y \frac{\partial^2 c}{\partial y^2} c_m(x, z) \cos(\lambda_m y)
$$

The z-derivative also follows the same process, which will be omitted for brevity. To put the equations into the form specified above, we evaluate 
the derivatives and then move over the advection terms. Doing this yields, as shown briefly above,

$$
    [\frac{\partial}{\partial x} [K_x \frac{\partial c(x, z)}{\partial x}] + \frac{\partial}{\partial z} [K_z \frac{\partial c(x, z)}{\partial z}] - u \frac{\partial c(x, z)}{\partial x} - w \frac{\partial c(x, z)}{\partial z}] Y_m(y)
$$

The rest of the terms in the equation come from the chain rule affects of the y derivative. The first term can be reorganized as

$$
    c_m(x, z) K_y' Y_m'(y)
$$

and the second term as

$$
    c_m(x, z) K_y Y_m''(y)
$$

where

$$
    Y_m''(y) = -\lambda_m^2 \cos(\lambda_m y) = -\lambda_m^2 Y_m(y)
$$

Substituting, we get

$$
    -\lambda_m^2 c_m(x, z) K_y Y_m(y)
$$

We have now completed most of the tedious algebra. We then can apply the operator $\int_0^{L_y} Y_n(y) dy$ to obtain the result

\begin{equation}
\begin{aligned}
    \sum_{m=0}^{\infty} \Bigg[ &\frac{\partial}{\partial x} \left( K_x \frac{\partial c(x, z)}{\partial x} \right) + \frac{\partial}{\partial z} \left( K_z \frac{\partial c(x, z)}{\partial z} \right) \\
    &- u \frac{\partial c(x, z)}{\partial x} - w \frac{\partial c(x, z)}{\partial z}  \alpha_{mn} \\
    &- v c_m(x, z) \beta_{mn} \\
    &- \lambda_m^2 c_m(x, z) \gamma_{mn} \\
    &+ c_m(x, z) \eta_{mn} \Bigg]
\end{aligned}
\end{equation}

We now will make some assumptions to simplify this equation. In the ABL, it is common to treat the speeds $v$ and $w$ as zero, such as in Tirabassi (1989).
We also assume that the force of advection is much more dominant than the force of diffusion in the x-direction. Mathematically, this is

$$
    u \frac{\partial c(x, z)}{\partial x} \gg \frac{\partial}{\partial x} (K_x\frac{\partial c(x, z)}{\partial x})
$$

Because of this, we can neglect the diffusion component $K_x$. We also consider that $K_y$ only depends on the z-direction.

We can now rewrite equation (11) above as:

\begin{equation}
    u \frac{\partial c(x, z)}{\partial x} = \frac{\partial} {\partial z} (K_z \frac{\partial c(x, z)}{\partial z}) - \lambda_m^2 K_y c_m(x, z)
\end{equation}

\subsection{An interpretation of this result}
